---
title: "My first markdown"
author: "Sigrid"
date: "9/19/2019"
output: html_document
---
## How to make a header
this is just text, R will not read it as code.
command + alt + i for creating a chunk (where you can code)

```{r setup}
library(tidyverse)
#Command 1 
knitr::opts_chunk$set(echo = TRUE)

#Command 2: set a new working directory to ALL chunks - not just the current chunk
#You need to remove the # for it to work
#knitr::opts_knit$set(root.dir = 'relative_path_to_root_from_Rmd' )

#Importing 

personality_test <- read_csv("NEW_CogSciPersonalityTest2019.csv")

```

setwd doesn't work in a r markdown fil! Save the rmd file in a folder, that you want to be the working directory. Just keep everything in your project. 

##Make a new chunk with an output

```{r}
df <- data.frame(personality_test)
#find class of balloon_balance
class(df$balloon_balance)
#R tells me it is numeric
```

### PART TWO

### summarise() function
This function collapses the whole dataframe into a single summary. For this function to work, it has to follow certain pattern.  First, you specify the data frame you want to summarise and then you say what values you want to have in your summary.

```{r}
library(tidyverse)

#Make a summary with just one value - the average shoesize 
summarise(df,mean(choose_rand_num))

#summary with several values: the average shoesize and its standard deviation
summarise(df,mean(shoesize),sd(shoesize))

```

summarise() is quite useless by itself, but everything changes when we **group** our data!

### group_by() function
group_by() takes an existing data frame and converts it into a data frame grouped by some principle. See examples in the chunk below:

### group_by function

##  Ggplot2

```{r}
ggplot(df)

#Examples of setups with different aesthetics:


ggplot(df, aes(x=df$gender))
ggplot(df, aes(x=df$gender, y=df$balloon))
ggplot(df, aes(x=breath_hold, color=gender))
```

#Geoms

```{r}
ggplot(df, aes(x = tongue_twist, fill = df$gender)) + geom_histogram(binwidth = 1)
```

#Make another histogram choose_rand_num
#To see all colors available colors()

```{r}

ggplot(df, aes(x = gender, y = df$choose_rand_num, fill = df$gender)) + geom_histogram(binwidth = 30)

```

###geom_errorbar

```{r}

ggplot(df, aes(x=gender, y=balloon, fill = gender))+
  geom_bar(stat='summary', fun.y = mean, fill = c("yellow2","skyblue3")) +
  geom_errorbar(stat = 'summary', fun.data = mean_se)

```

### More layers and aesthetics 

Explanation of each of the geoms:
- fill = gender == fill the bars in different colors according to the values female/male = gender
- geom_bar == how to get the data into the plot
-- width = width of the bar (notice the dot insted of comma)
- geom_errorbar == shows the standard error of the mean 
- labs = the name of the values of the x- and y-axis respectively 
- theme_minimal() == white background instead of grey 

```{r}
ggplot(df, aes(x=gender, y=balloon, fill = gender))+
  geom_bar(stat='summary', fun.y = mean, width = 0.5)+
  geom_errorbar(stat = 'summary', fun.data = mean_se, width = 0.2)+labs(x = "Gender", y = "Balloon time")+
  theme_minimal()
  

```

#### Visualization exercise
1. Identically to bar plots we've made above, make a bar plot showing average shoesize according to handedness.

2. Try to make it pretty!

```{r}
ggplot(df, aes(x = df$handedness, y = mean(df$shoesize), fill = df$handedness)) + labs(x = "Handedness", y = "Shoesize (mean)") + theme_minimal() + geom_bar(stat='summary', fun.y = mean, width = 0.5)



#+ geom_histogram(binwidth = 10) + geom_bar(stat="count",fun.y = mean)
```

Ser ikke helt rigtigt ud... 


