---
title: "My first markdown"
author: "Sigrid"
date: "9/19/2019"
output: html_document
---
## How to make a header
this is just text, R will not read it as code.
command + alt + i for creating a chunk (where you can code)

```{r setup}
library(tidyverse)
#Command 1 
knitr::opts_chunk$set(echo = TRUE)

#Command 2: set a new working directory to ALL chunks - not just the current chunk
#You need to remove the # for it to work
#knitr::opts_knit$set(root.dir = 'relative_path_to_root_from_Rmd' )

#Importing 

personality_test <- read_csv("NEW_CogSciPersonalityTest2019.csv")

```

setwd doesn't work in a r markdown fil! Save the rmd file in a folder, that you want to be the working directory. Just keep everything in your project. 

##Make a new chunk with an output

```{r}
df <- data.frame(personality_test)
#find class of balloon_balance
class(df$balloon_balance)
#R tells me it is numeric
```

### PART TWO

### summarise() function
This function collapses the whole dataframe into a single summary. For this function to work, it has to follow certain pattern.  First, you specify the data frame you want to summarise and then you say what values you want to have in your summary.

```{r}
library(tidyverse)

#Make a summary with just one value - the average shoesize 
summarise(df,mean(choose_rand_num))

#summary with several values: the average shoesize and its standard deviation
summarise(df,mean(shoesize),sd(shoesize))

```

summarise() is quite useless by itself, but everything changes when we **group** our data!

### group_by() function
group_by() takes an existing data frame and converts it into a data frame grouped by some principle. See examples in the chunk below:

### group_by function
