---
title: "2019-11-20 regression 3"
author: "Sigrid"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, pastecs, lme4)

```

## Winter (2014) Tutorial 2

```{r importing data}

politeness <- read_csv("politeness_data.csv")

```


Following the tutorial doing regression. 

```{r}

head(politeness)
tail(politeness)
summary(politeness)
str(politeness)
colnames(politeness)

which(is.na(politeness$frequency))

```

```{r boxplot}
boxplot(frequency ~ attitude*gender, col=c("white", "lightgray"), politeness)

```

```{r model}

politeness.model <-  lmer(frequency ~ attitude + gender + (1|subject) + (1|scenario), politeness)
summary(politeness.model)
```

```{r significance}

politeness.null <- lmer(frequency ~ gender + (1|subject) + (1|scenario), politeness, REML = F)

politeness.model <-  lmer(frequency ~ attitude + gender + (1|subject) + (1|scenario), politeness, REML = F)

anova(politeness.null, politeness.model)

```

Politeness affected pitch (chi i anden (1) = 11.62, p = 0.00065), lowering it by about 19.7 Hz +- 5.6 (standard errors)

hiker analogy: you can only drop one piece of luggage at the time to know what made a difference. 

### interaction
full model = frequency ~ attitude*gender
reduced model = frequency ~ attitude + gender
theory = if attitude is different for men and women (for attitude to have an affect on pitch (frequency) it is somehow modulated through gender)
interactions are specified with "*" 

# Random slopes versus random intercepts

Random intercept model = accounts for baseline-differences in frequency (ptich)
might not be a valid assumption, we need a model that can make different slopes for the different subjects. 
```{r random intercept model}

coef(politeness.model)

```

```{r random slopes}
politeness.model <- lmer(frequency ~attitude + gender + (1+ attitude|subject) + (1 + attitude | scenario), politeness, REML = F)

coef(politeness.model)

politeness.null <- lmer(frequency ~ gender + (1 + attitude|subject) + (1+attitude|scenario), politeness, REML = F)

anova(politeness.null, politeness.model)

```


